author: "H2030 Genome Center DSP/DAIP"
output:
  html_document: default
  word_document: default
header-includes: \usepackage{graphicx}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, echo=FALSE, include=FALSE}
library(ggplot2)
library(kableExtra)
library(ggpubr)
library(scales)
```


```{r}
raw=read.table("summaryR4.csv",sep=",",header=T)
frm=data.frame(raw)
frm$sample = reorder(frm$sample,frm$coverage)
frm$quality = round(frm$human_bait_reads/(frm$tot_mapreads-frm$numreads + 1), digits=2) 
```

```{r}
frm2=frm[c("sample","tot_numreads","tot_mapreads")]
frm2$human_on_target_reads=frm$human_bait_reads
frm2$SARScov2_reads=frm$numreads
frm2$SARScov2_coverage=frm$coverage
frm2$SARScov2_meandepth=frm$meandepth
frm2$perpos15x=frm$perpos15x
frm2$perN5=frm$perN5
frm2$human_on_target_proportion=frm$quality
frm2[c("tot_numreads","tot_mapreads","human_on_target_reads","SARScov2_reads")] = frm2[c("tot_numreads","tot_mapreads","human_on_target_reads","SARScov2_reads")]/1e6
```

```{r}
frm3=frm2[frm2$SARScov2_meandepth < 2000,]
```

## General observations

Total number of samples:
```{r}
nrow(raw)
```

```{r}
frm2 %>%
    kable() %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

## Samples with mean depth < 2000

Total number of samples:
```{r}
nrow(frm3)
```

```{r}
frm3 %>%
    kable() %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

# Read counts category distribution

```{r}
quality_cutoff=0.5

frm$ControlNeg=grepl("CtlNeg",frm$sample,fixed=TRUE) 
frm$ControlPos=grepl("CtlPos",frm$sample,fixed=TRUE) 

frm$control='Sample'
noGp=nrow(frm[frm$perpos15x < 95 & frm$quality < quality_cutoff,])
noGg=nrow(frm[frm$perpos15x < 95 & frm$quality >= quality_cutoff,])
yesGp=nrow(frm[frm$perpos15x >= 95 & frm$quality < quality_cutoff,])
yesGg=nrow(frm[frm$perpos15x >= 95 & frm$quality >= quality_cutoff,])
neg=nrow(frm[frm$ControlNeg,])
pos=nrow(frm[frm$ControlPos,])
p15plus=nrow(frm[frm$perpos15x >=95,])
p15minus=nrow(frm[frm$perpos15x < 95,])

frm$quadrant=0
if (yesGg>0) {
    frm[frm$perpos15x >= 95 & frm$quality >= quality_cutoff,]$quadrant=paste(yesGg,"pass GISAID / good quality",sep=" ")
}
if (yesGp>0) {
    frm[frm$perpos15x >= 95 & frm$quality < quality_cutoff,]$quadrant=paste(yesGp,"pass GISAID / poor quality",sep=" ")
}
if (noGg>0) {
    frm[frm$perpos15x < 95 & frm$quality >= quality_cutoff,]$quadrant=paste(noGg,"not passing GISAID / good quality",sep=" ")
}
if (noGp>0) {
    frm[frm$perpos15x < 95 & frm$quality < quality_cutoff,]$quadrant=paste(noGp,"not passing GISAID / poor quality",sep=" ")
}
if (pos>0) {
      frm[frm$ControlPos,]$quadrant=paste(pos, "positive controls", sep=" ")
}
if (neg>0) {
      frm[frm$ControlNeg,]$quadrant=paste(neg, "negative controls", sep=" ")
}
```

#### Number of SARS-CoV-2 reads in respect to proportion of human reads on target

```{r}
p<-ggplot(frm, aes(x=quality,y=numreads,colour=quadrant)) +
    geom_point(aes(shape=quadrant),alpha=0.8) +
    theme() +
    xlab("Proportion of human reads on target") +
    ylab("SARS-CoV-2 reads") +
    scale_shape_manual(values=c(1,8,3,17,24,9)) +
    scale_color_manual(values=c("green", "red", "blue","orange","brown","black")) +
    scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),breaks = c(0,10,100,1000,10000,100000,1000000,10000000,100000000), labels = trans_format("log10", math_format(10^.x))) +
    xlim(0,1)
p<-p+theme_bw()
p<-p+theme(legend.title = element_blank())
print(p)
```

#### Number of SARS-CoV-2 reads in respect to number of human reads on target


```{r}
p<-ggplot(frm, aes(x=human_bait_reads,y=numreads,colour=quadrant)) +
    geom_point(aes(shape=quadrant),alpha=0.8) +
    theme() +
    xlab("Human reads on target") +
    ylab("SARS-CoV-2 reads") +
    scale_shape_manual(values=c(1,8,3,17,24,9)) +
    scale_color_manual(values=c("green", "red", "blue","orange","brown","black")) +
    scale_x_continuous(trans=scales::pseudo_log_trans(base = 10),breaks = c(0,10,100,1000,10000,100000,1000000,10000000,100000000), labels = trans_format("log10", math_format(10^.x))) +
    scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),breaks = c(0,10,100,1000,10000,100000,1000000,10000000,100000000), labels = trans_format("log10", math_format(10^.x)))
p<-p+theme_bw()
p<-p+theme(legend.title = element_blank())
print(p)
```

```{r}
p<-ggplot(frm, aes(x=human_bait_reads,y=numreads,colour=quadrant)) +
    geom_point(aes(shape=quadrant),alpha=0.5) +
    theme() +
    xlab("Human reads on target") +
    ylab("SARS-CoV-2 reads") +
    scale_shape_manual(values=c(1,8,3,17,24,9)) +
    scale_color_manual(values=c("green", "red", "blue","orange","brown","black")) +
    scale_x_continuous(trans=scales::pseudo_log_trans(base = 10),breaks = c(0,10,100,1000,10000,100000,1000000,10000000,100000000), labels = trans_format("log10", math_format(10^.x))) +
    scale_y_continuous(trans=scales::pseudo_log_trans(base = 10),breaks = c(0,10,100,1000,10000,100000,1000000,10000000,100000000), labels = trans_format("log10", math_format(10^.x))) +
    facet_wrap("quadrant",ncol=2) +
    theme(legend.position = "none")

p<-p+theme_bw()
p<-p+theme(legend.title = element_blank())
print(p)
```


# Number of reads

```{r}
p<-ggplot(frm, aes(x=sample,y=tot_numreads,colour="Total")) +
        geom_point(alpha=0.6) +
        geom_point(aes(x=sample,y=tot_mapreads,colour="Mapped"),alpha=0.6) +
        geom_point(aes(x=sample,y=numreads,colour="SARS-CoV2"),alpha=0.6) +
        geom_point(aes(x=sample,y=human_bait_reads,colour="Human baits"),alpha=0.6) +
        xlab("Samples (ordered by increasing coverage)") +
        ylab("Num reads")+
        scale_colour_manual("",breaks = c("Total","Mapped","SARS-CoV2","Human baits"), values = c("blue","green","red","yellow"))

p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```

# Percentage of covered positions

```{r}
p<-ggplot(frm, aes(x=sample,y=coverage)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions covered")+
        ylim(0,100)

p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```

# Mean coverage depth

```{r}
p<-ggplot(frm, aes(x=sample,y=meandepth)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("Mean coverage depth")

p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```

# Percentage of positions achieving 15x coverage depth

```{r}
p<-ggplot(frm, aes(x=sample,y=perpos15x)) +
        geom_point(alpha=0.6) +
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions >= 15x")+
        ylim(0,100)
p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```


# Percentage of positions achieving 50x coverage depth

```{r}
p<-ggplot(frm, aes(x=sample,y=perpos50x)) +
        geom_point(alpha=0.6) +
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions >= 50x")+
        ylim(0,100)
p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```


# Percentage of positions with less than 5x coverage

```{r}
p<-ggplot(frm, aes(x=sample,y=perN5)) +
        geom_point(alpha=0.6) +
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions < 5x")+
        ylim(0,100)
p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```

# Number of human reads on target

```{r}
p<-ggplot(frm, aes(x=sample,y=quality)) +
        geom_point(alpha=0.6) +
        xlab("Samples (ordered by increasing SARS-CoV-2 coverage)") +
        ylab("Number of human reads on target")+
        ylim(0,1)
p<-p+theme_bw()
p<-p+theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
p<-p+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
print(p)
```


