author: "H2030 Genome Center DSP/DAIP"
output:
  html_document: default
  word_document: default
header-includes: \usepackage{graphicx}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r, echo=FALSE, include=FALSE}
library(ggplot2)
library(kableExtra)
options(digits=11)
```


```{r}
raw=read.table("summaryR4.csv",sep=",",header=T)
frm=data.frame(raw)
frm$sample = reorder(frm$sample,frm$coverage)
```

## General observations

Total number of samples:
```{r}
nrow(raw)
```

```{r}
frm %>%
    kable() %>%
    kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
```

# Read counts

```{r}
p<-ggplot(frm, aes(x=sample,y=tot_numreads,colour="Total")) +
        geom_point(alpha=0.6) +
        geom_point(aes(x=sample,y=tot_mapreads,colour="SARS-CoV2/Human"),alpha=0.6) +
        geom_point(aes(x=sample,y=numreads,colour="SARS-CoV2"),alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("Num reads")+
        scale_colour_manual("",breaks = c("Total","SARS-CoV2/Human","SARS-CoV2"), values = c("blue","green","red"))
print(p)
```

# Percentage of covered positions

```{r}
p<-ggplot(frm, aes(x=sample,y=coverage)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions covered")+
        ylim(0,100)
print(p)
```

# Mean coverage depth

```{r}
p<-ggplot(frm, aes(x=sample,y=meandepth)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("Mean coverage depth")
print(p)
```

# Percentage of positions achieving 15x coverage depth

```{r}
p<-ggplot(frm, aes(x=sample,y=perpos15x)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions >= 15x")+
        ylim(0,100)
print(p)
```


# Percentage of positions achieving 50x coverage depth

```{r}
p<-ggplot(frm, aes(x=sample,y=perpos50x)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions >= 50x")+
        ylim(0,100)
print(p)
```


# Percentage of positions with less than 5x coverage

```{r}
p<-ggplot(frm, aes(x=sample,y=perN5)) +
        geom_point(alpha=0.6) +
        theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())+
        xlab("Samples (ordered by increasing coverage)") +
        ylab("%positions < 5x")+
        ylim(0,100)
print(p)
```

